---
title: "Figure 2"
author: "Alexandra G. Mitchell and Francesca Fardo"
date: "28/02/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load required packages
if (!require("pacman")) install.packages("pacman")
  pacman::p_load(gghalves, ggpol, ggpubr, tidyverse,
                 gridExtra, wesanderson, RColorBrewer, Rmisc)
```
 
##### FIGURES
```{r}
# threshold summary data
filename <- file.path('data', 'summary_tsl2_208.csv')
if (file.exists(filename)){
  df <- read.csv(filename)
} else {
  print('Data file does not exist, run tsl2_phs_q2.Rmd first')
}

# model data 
conN_file <- file.path('data', 'model_contrast_data.csv')
if (file.exists(conN_file)){
  df_conN <- read.csv(conN_file)
} else {
  print('Data file does not exist, run tsl2_phs_q3.Rmd first')
}

# constrast summary data
sum_file <- file.path('data', 'con_summary_208.csv')
if (file.exists(sum_file)){
  con_cond <- read.csv(sum_file)
} else {
  print('Data file does not exist, run tsl2_phs_q3.Rmd first')
}

# statistics
stats_file <- file.path('results', 'phs_contrast_stats.csv')
if (file.exists(stats_file)){
  con_stats <- read.csv(stats_file)
} else {
  print('Data file does not exist, run tsl2_phs_q3.Rmd first')
}

# Define colour palettes 
blue <- brewer.pal(9, "Blues")
color_light <- blue[4]
color_dark <- blue[7]
greys9  <- brewer.pal(9, "Greys")
blues5 <- brewer.pal(5, "Blues")
greens9  <- brewer.pal(9, "Greens")
orans9  <- brewer.pal(9, "Oranges")
dj2 <- wes_palettes$Darjeeling2

# Define figure features
axistext <- 10
axistitle <- 10
title <- 12
legend <- 9

# Define plot nudges
nudge1 <- 0.20
nudge2 <- 0.30
nudge3 <- -0.12
nudge4 <- 0.12
nudge5 <- -.20
nudge6 <- -.3
```

# Panel 1: Threshold by starting temperature
```{r panel1}
# Select data for histogram
df_phs_fig <- df %>% 
  dplyr::filter(phs != 0)

# Rename instruction
df_phs_fig[which(df_phs_fig$instruction == "detect"), ]$instruction <- "Innocuous"
df_phs_fig[which(df_phs_fig$instruction == "pain"), ]$instruction <- "Noxious"

# Create plot
p1 <- ggplot(data = df_phs_fig, 
             aes(x = baseline, 
                 color = instruction, 
                 fill = forcats::fct_rev(factor(phs)))) +
  geom_histogram(width = 3, stat = "count", position = "stack", alpha = .8) +
  facet_wrap(~ instruction, scales = 'free') +
  scale_fill_manual(values = c(orans9[9],orans9[6],orans9[3])) + 
  scale_color_manual(values = c('black', 'black')) + 
  labs(x=NULL, 
       y="PHS-responders") +
  guides(fill=guide_legend(title="Number of PHS")) +
  theme_classic() +
  theme(panel.grid.minor = element_blank(), # text size for labels on the x-axis
        strip.text = element_text(size = 10), # text size and color on facet_wrap        
        #strip.background =element_rect(fill=c(blue5[5],blue5[5])),
        axis.text = element_text(size = axistext),
        axis.title = element_text(size = axistitle),
        legend.position = 'bottom',
        #legend.background = element_rect(colour = "gray", fill = "white", linetype="solid"),
        legend.text = element_text(size = legend),
        legend.title = element_text(size = legend)) +
  guides(color = FALSE)  +
  scale_x_continuous(breaks = c(44,38,32), labels=c("High", "Medium", "Low"), ) +#, limits = c(44,38,32)) +
  ylim(0,60)

p1 <- annotate_figure(p1, top = text_grob('A. Effect of Contrast on PHS',
                      size = 11, face = 'bold'))

p1

ggsave(here::here("figures", "F2_panel1.png"), plot = p1, 
       device = NULL, width = 8, height = 5, dpi = 600)
```
# Panel 2: TSL thresholds 
Plot the change in pain and detection threshold over starting temperature using rainclouds
```{r panel3}
plot_df <- df %>% 
  mutate(baseline = replace(baseline, baseline == 32, 1)) %>% 
  mutate(baseline = replace(baseline, baseline == 38, 2)) %>% 
  mutate(baseline = replace(baseline, baseline == 44, 3))
# Rename instruction
plot_df[which(plot_df$instruction == "detect"), ]$instruction <- "Innocuous"
plot_df[which(plot_df$instruction == "pain"), ]$instruction <- "Noxious"

# Jitter
plot_df$xj <- jitter(plot_df$baseline, amount = .05)

# for this plot, remove trials that are 0 or < 0, for illustrative purposes
plot_df <- plot_df %>% 
  dplyr::filter(threshold > 0.1)
```

# Calculate group-level summary of average innocuous and noxious TSL thresholds for rainclouds
```{r panel2_stats}
# Define total number of participants
N = length(unique(df$subject))

# Define summary stats to compute
my_funcs <- list(
  mean = mean, 
  median = median, 
  sd = sd, 
  stderr = function(x) sd(x)/sqrt(N),
  ci_upper = function(x) CI(x, ci = 0.95)[1],
  ci_mean = function(x) CI(x, ci = 0.95)[2],
  ci_lower = function(x) CI(x, ci = 0.95)[3])

# Get group-level summary stats
tsl2_group_stats <- plot_df %>%
  group_by(baseline, instruction) %>%
  dplyr::reframe(across(threshold, my_funcs)) %>%
  dplyr::rename(mean = threshold_mean,
                median = threshold_median,
                sd = threshold_sd,
                stderr = threshold_stderr,
                ci_upper = threshold_ci_upper,
                ci_lower = threshold_ci_lower,
                ci_mean = threshold_ci_mean) %>%
  mutate(ci_diff = ci_upper - ci_lower,
         ci = ci_upper - ci_mean)

tsl2_group_stats

# Recode baseline in group-level summary
tsl2_group_stats <- tsl2_group_stats %>% 
  mutate(baseline = replace(baseline, baseline == 32, 1)) %>% 
  mutate(baseline = replace(baseline, baseline == 38, 2)) %>% 
  mutate(baseline = replace(baseline, baseline == 44, 3))
# Rename instruction
tsl2_group_stats[which(tsl2_group_stats$instruction == "detect"), ]$instruction <- "Innocuous"
tsl2_group_stats[which(tsl2_group_stats$instruction == "pain"), ]$instruction <- "Noxious"

```

```{r panel3}
# Plot 2
p3 <- ggplot(data = plot_df, 
             mapping = aes(x = baseline, y = threshold,
                           group = instruction, colour = instruction, fill = instruction)) +
  # Add individual points
  geom_point(aes(x = xj), shape = 21, size = .8, alpha = .6,
             position = position_nudge(x = nudge3)) +
  facet_wrap(~instruction, scales = 'free') +
  geom_boxplot(
    data = plot_df %>% 
      dplyr::filter(baseline == 1), outlier.color = NA,
    position = position_nudge(x = nudge4), 
    width = .18,  notch = TRUE, lwd = .5, fatten = 1) +
  geom_boxplot(
    data = plot_df %>% 
      dplyr::filter(baseline == 2), outlier.color = NA,
    position = position_nudge(x = nudge4), 
    width = .18,  notch = TRUE, lwd = .5, fatten = 1) +
  geom_boxplot(
    data = plot_df %>% 
      dplyr::filter(baseline == 3), outlier.color = NA,
    position = position_nudge(x = nudge4), 
    width = .18,  notch = TRUE, lwd = .5, fatten = 1) +
  
  # Add a line connecting the two means (Innocuous TSL)
  geom_line(data = tsl2_group_stats,
            aes(x = c(baseline + nudge4), y = median), size = .5) +
  
  # Define additional settings
  scale_x_continuous(breaks = c(1, 2, 3), labels=c("Low", "Medium", "High"), ) +
  scale_colour_manual(labels = c("Innocuous", "Noxious"), values=c(blues9[7],blues9[8])) +
  scale_fill_manual(labels = c("Innocuous", "Noxious"), values=c(blues9[2],blues9[3])) +
  labs(y = "TSL Threshold (ºC)", x = NULL) +
  theme_classic() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(), # text size for labels on the x-axis
        strip.text = element_text(size = 10), # text size and color on facet_wrap 
        axis.text = element_text(size = axistext),
        axis.title = element_text(size = axistitle),
        legend.text = element_text(size = legend),
        legend.title = element_blank()) # remove legend


# add title
p3 <- annotate_figure(p3, top = text_grob('B. Individual TSL Thresholds',
                      size = 11, face = 'bold'))

p3

#ggsave('Thresholds.png', p2, path = Apath, device = NULL, width = 7, height = 4.5, dpi = 1000)
ggsave(here::here("figures", "F2_panel2.png"), plot = p3, 
       device = NULL, width = 8, height = 5, dpi = 600)
```



# Make something similar for contrast
# Contrast by PHS
```{r}
# calculate and add pain contrast into summary
pain_con <- con_cond %>% 
  dplyr::filter(instruction == 'pain') %>% 
  dplyr::rename(TCFpain = TCF,
                threshold_pain = threshold) %>% 
  dplyr::select(c(subject,baseline,threshold_pain,TCFpain))
  
detect_con <- con_cond %>% 
  dplyr::filter(instruction == 'detect') %>% 
  dplyr::rename(TCFdetect = TCF,
                threshold_detect = threshold)

plot_con <- merge(detect_con, pain_con, by = c('subject','baseline'))

# Recode baseline
plot_con <- plot_con %>% 
  mutate(baseline = replace(baseline, baseline == 32, 1)) %>% 
  mutate(baseline = replace(baseline, baseline == 38, 2)) %>% 
  mutate(baseline = replace(baseline, baseline == 44, 3)) %>% 
  select(-c(instruction)) %>% 
  # lengthen data-frame by TCF to facet wrap
  pivot_longer(cols = c(TCFdetect, TCFpain), names_to = 'instruction', values_to = 'TCF')

# instruction names
plot_con[which(plot_con$instruction == "TCFdetect"), ]$instruction <- "Innocuous"
plot_con[which(plot_con$instruction == "TCFpain"), ]$instruction <- "Noxious"

# Jitter
plot_con$xj <- jitter(plot_con$baseline, amount = .04)
plot_con$phs_01 <- as.factor(plot_con$phs_01)
#plot_con$baseline <- as.numeric(as.character(plot_con$baseline))
```

# Contrast summary statistics
```{r}
# Get group-level summary stats
con_group_stats <- plot_con %>%
  group_by(baseline, instruction, phs_01) %>%
  dplyr::reframe(across(TCF, my_funcs)) %>%
  dplyr::rename(mean = TCF_mean,
                median = TCF_median,
                sd = TCF_sd,
                stderr = TCF_stderr,
                ci_upper = TCF_ci_upper,
                ci_lower = TCF_ci_lower,
                ci_mean = TCF_ci_mean) %>%
  mutate(ci_diff = ci_upper - ci_lower,
         ci = ci_upper - ci_mean)

con_group_stats
```

# Panel 3:  phs detection thresolds
```{r panel3}
# function for making labels 1dp
scaleFUN <- function(x) sprintf("%.1f", x)

p4 <- ggplot(data = plot_con, 
             mapping = aes(x = baseline, 
                           y = TCF, 
                           group = as.factor(phs_01), 
                           fill = as.factor(phs_01), 
                           colour = as.factor(phs_01),
                           alpha = as.factor(instruction))) +
  facet_wrap(~instruction, scales = 'free') +
  geom_boxplot(
    data = plot_con %>% 
      filter(baseline == 1),
    position = position_dodge(.5), width = .28, notch = FALSE,
    lwd = .4, fatten = .8, outlier.shape = 21, outlier.size = .7) +
  geom_boxplot(
    data = plot_con %>% 
      filter(baseline == 2),
    position = position_dodge(.5), width = .28, notch = FALSE,
    lwd = .4, fatten = .8, outlier.shape = 21, outlier.size = .7) +
  geom_boxplot(
    data = plot_con %>% 
      filter(baseline == 3),
    position = position_dodge(.5), width = .28, notch = FALSE,
    lwd = .4, fatten = .8, outlier.shape = 21, outlier.size = .7) +
  scale_fill_manual(labels = c("No PHS", "PHS"), values=c(blues9[3],orans9[3])) +
  scale_colour_manual(labels = c("No PHS", "PHS"), values=c(blues9[7],orans9[7])) +
  scale_alpha_manual(values = c(.5,.9), guide = 'none') +
  # Define additional settings
  scale_x_continuous(breaks=c(1,2,3), labels = c('Low','Medium','High')) +
  scale_y_continuous(labels=scaleFUN) +
  labs(y = "Contrast (Tmax - Tmin)/50",
       x = "Contrast Condition") +
  #ylim(0,0.55) +
  theme_classic() +
  theme(legend.position = 'bottom',
        legend.title = element_blank(),
        panel.grid.minor = element_blank(), # text size for labels on the x-axis
        strip.text = element_text(size = 10), # text size and color on facet_wrap
        plot.title = element_text(hjust = 0.5, size = title),
        axis.text = element_text(size = axistext),
        axis.title = element_text(size = axistitle),
        legend.text = element_text(size = legend)
        )

# add title
p4 <- annotate_figure(p4, top = text_grob('C. Thermal Contrast by PHS',
                      size = 11, face = 'bold'))
p4
ggsave(here::here("figures", "F2_panel3.png"), plot = p4, 
       device = NULL, width = 9, height = 4.5, dpi = 600)

```

# Put all panels together
```{r figure1}
# setting placement of figures
F2 <- ggarrange(p1, NA, p3, NA, p4,
                 ncol = 1, nrow = 5,
                 heights = c(.95,.07,.78,.07,1))

F2

ggsave(file.path("figures", "Figure2.png"), 
       plot = F2, device = NULL, width = 3.5, height = 9, dpi = 600)
```
