---
title: "Figure 3"
author: "A.G. Mitchell & F. Fardo"
date: "2023-05-02"
output: pdf_document
---

NOTE: THE DATA FOR THE MODEL SIMULATIONS: m*-bootstrapAUCs.csv ARE TOO LARGE TO STORE ON GITHUB. 
THEY ARE THEREFORE AVAILABLE FOR DOWNLOAD FROM THIS OSF LINK: [ADD]

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load required packages
if (!require("pacman")) install.packages("pacman")
  pacman::p_load(cowplot, effects, gghalves, ggpol, ggpubr, gridExtra, wesanderson, 
                 RColorBrewer, tidyverse)
```

A reminder of the models: 
1: phs~baseline
2: phs~TCFdetect
3: phs~TCFpain
4: phs~TCFdetect * TCFpain
5: phs~TCFdetect * TCFnback
6: phs~TCFdetect * TCFpain * TCFnback
7: phs~threshold_detect * threshold_pain

##### FIGURE 3 CODE
# Define paths, load files and models
# make sure models are in current working directory under a folder called results
```{r}
# load models
m1_file <- file.path("results", "h3_mod1.RData")
m2_file <- file.path("results", "h3_mod2.RData")
m3_file <- file.path("results", "h3_mod3.RData")
m4_file <- file.path("results", "h3_mod4.RData")
m5_file <- file.path("results", "h3_mod5.RData")
m6_file <- file.path("results", "h3_mod6.RData")
m7_file <- file.path("results", "h3_mod7.RData")

if (file.exists(m1_file)){
  mod.m1 <- readRDS(m1_file)
  mod.m2 <- readRDS(m2_file)
  mod.m3 <- readRDS(m3_file)
  mod.m4 <- readRDS(m4_file)
  mod.m5 <- readRDS(m5_file)
  mod.m6 <- readRDS(m6_file)
  mod.m7 <- readRDS(m7_file)
} else {
  print('Models do not exist, run tsl2_phs_q3.Rmd and check current directory')
}

# load model fits
m1_ROC_file <- file.path("results", "h3_model1_ROC.csv")
m2_ROC_file <- file.path("results", "h3_model2_ROC.csv")
m3_ROC_file <- file.path("results", "h3_model3_ROC.csv")
m4_ROC_file <- file.path("results", "h3_model4_ROC.csv")
m5_ROC_file <- file.path("results", "h3_model5_ROC.csv")
m6_ROC_file <- file.path("results", "h3_model6_ROC.csv")
m7_ROC_file <- file.path("results", "h3_model7_ROC.csv")
if (file.exists(m1_ROC_file)){
  m1_ROC <- read.csv(m1_ROC_file)
  m2_ROC <- read.csv(m2_ROC_file)
  m3_ROC <- read.csv(m3_ROC_file)
  m4_ROC <- read.csv(m4_ROC_file)
  m5_ROC <- read.csv(m5_ROC_file)
  m6_ROC <- read.csv(m6_ROC_file)
  m7_ROC <- read.csv(m7_ROC_file)
} else {
  print('Models ROCs do not exist, run tsl2_phs_q3.Rmd and check current directory')
}

# finally, load bootstrapped data
  # location of .csv files (can get them from linked OSF page)
Apath = '/Users/au706616/Documents/Experiments/PHS-SKULDNET/analyses/h3' 
# filenames
m1_bsfile <- file.path(Apath, 'm1-bootstrapAUCs.csv')
m2_bsfile <- file.path(Apath, 'm2-bootstrapAUCs.csv')
m3_bsfile <- file.path(Apath, 'm3-bootstrapAUCs.csv')
m4_bsfile <- file.path(Apath, 'm4-bootstrapAUCs.csv')
m5_bsfile <- file.path(Apath, 'm5-bootstrapAUCs.csv')
m6_bsfile <- file.path(Apath, 'm6-bootstrapAUCs.csv')
m7_bsfile <- file.path(Apath, 'm7-bootstrapAUCs.csv')
  # load file
if (file.exists(m1_bsfile)){
  m1_bs_perf <- read.csv(m1_bsfile)
  m2_bs_perf <- read.csv(m2_bsfile)
  m3_bs_perf <- read.csv(m3_bsfile)
  m4_bs_perf <- read.csv(m4_bsfile)
  m5_bs_perf <- read.csv(m5_bsfile)
  m6_bs_perf <- read.csv(m6_bsfile)
  m7_bs_perf <- read.csv(m7_bsfile)
} else {
  print('Data files do not exist. They can be downloaded from:   https://osf.io/t3n8s/')
}

# finally, model data
res_file <- file.path('data', 'model_contrast_data.csv')
if (file.exists(res_file)){
  df_conN <- read.csv(res_file)
} else {
  print('Data file does not exist, run tsl2_phs_q3.Rmd first')
}
```

# Figure parameters
```{r}
# colour palettes etc
greys9  <- brewer.pal(9, "Greys")
blues5 <- brewer.pal(5, "Blues")
grens9 <- brewer.pal(9, "Greens") 
orans9 <- brewer.pal(9, "Oranges") 
dj2 <- wes_palettes$Darjeeling2
color_light <- blues9[4]
color_dark <- blues9[7]

# figure features
axistext = 10
axistitle = 12
title = 12
legend = 8
# plot nudge
nudge1 = 0.20
nudge2 = 0.30
nudge3 = -0.15
nudge4 = 0.15
```

# FIGURE 3
# Plotting the AUC and CIs from simulated data
```{r}
# calculate 95% ci quant val
v = qnorm(.975)
N = 208

# first remove models with warnings - both data-frames
# model 1
#m1_bs_perf = m1_bs_perf[is.na(m1_bs_perf$warnings) ,]
m1_bs_perf1 = m1_bs_perf[is.finite(m1_bs_perf$Cutoff) ,] # get rid of infinite values, for means
# average AUC and confidence intervals - simple model
m1_perf <- aggregate(Cutoff~model*Area_under_the_ROC_curve, 
                      mean, data = m1_bs_perf1)
# then calculate mean and CI of the boostrap
m1_AUC_mean <- mean(m1_perf$Area_under_the_ROC_curve)
m1_AUC_sd <- sd(m1_perf$Area_under_the_ROC_curve)
# CI from SD and N
m1_AUC_ci <- v*(m1_AUC_sd/sqrt(N))
m1_AUC_CIs <- cbind((m1_AUC_mean - m1_AUC_ci), (m1_AUC_mean + m1_AUC_ci))

# model 2
#m2_bs_perf = m2_bs_perf[is.na(m2_bs_perf$warnings) ,]
m2_bs_perf1 = m2_bs_perf[is.finite(m2_bs_perf$Cutoff) ,]
# average AUC and confidence intervals 
m2_perf <- aggregate(Cutoff~model*Area_under_the_ROC_curve, 
                      mean, data = m2_bs_perf1)
# then calculate mean and CI
m2_AUC_mean <- mean(m2_perf$Area_under_the_ROC_curve)
m2_AUC_sd <- sd(m2_perf$Area_under_the_ROC_curve)
# CI from SD and N
m2_AUC_ci <- v*(m2_AUC_sd/sqrt(N))
m2_AUC_CIs <- cbind((m2_AUC_mean - m2_AUC_ci), (m2_AUC_mean + m2_AUC_ci))

# model 3
#m3_bs_perf = m3_bs_perf[is.na(m3_bs_perf$warnings) ,]
m3_bs_perf1 = m3_bs_perf[is.finite(m3_bs_perf$Cutoff) ,]
# average AUC and confidence intervals 
m3_perf <- aggregate(Cutoff~model*Area_under_the_ROC_curve, 
                      mean, data = m3_bs_perf1)
# then calculate mean and CI
m3_AUC_mean <- mean(m3_perf$Area_under_the_ROC_curve)
m3_AUC_sd <- sd(m3_perf$Area_under_the_ROC_curve)
# CI from SD and N
m3_AUC_ci <- v*(m3_AUC_sd/sqrt(N))
m3_AUC_CIs <- cbind((m3_AUC_mean - m3_AUC_ci), (m3_AUC_mean + m3_AUC_ci))

# model 4
#m4_bs_perf = m4_bs_perf[is.na(m4_bs_perf$warnings) ,]
m4_bs_perf1 = m4_bs_perf[is.finite(m4_bs_perf$Cutoff) ,]
# average AUC and confidence intervals 
m4_perf <- aggregate(Cutoff~model*Area_under_the_ROC_curve, 
                      mean, data = m4_bs_perf1)
# then calculate mean and CI
m4_AUC_mean <- mean(m4_perf$Area_under_the_ROC_curve)
m4_AUC_sd <- sd(m4_perf$Area_under_the_ROC_curve)
# CI from SD and N
m4_AUC_ci <- v*(m4_AUC_sd/sqrt(N))
m4_AUC_CIs <- cbind((m4_AUC_mean - m4_AUC_ci), (m4_AUC_mean + m4_AUC_ci))

# model 5
#m5_bs_perf = m5_bs_perf[is.na(m5_bs_perf$warnings) ,]
m5_bs_perf1 = m5_bs_perf[is.finite(m5_bs_perf$Cutoff) ,]
# average AUC and confidence intervals 
m5_perf <- aggregate(Cutoff~model*Area_under_the_ROC_curve, 
                      mean, data = m5_bs_perf1)
# then calculate mean and CI
m5_AUC_mean <- mean(m5_perf$Area_under_the_ROC_curve)
m5_AUC_sd <- sd(m5_perf$Area_under_the_ROC_curve)
# CI from SD and N
m5_AUC_ci <- v*(m5_AUC_sd/sqrt(N))
m5_AUC_CIs <- cbind((m5_AUC_mean - m5_AUC_ci), (m5_AUC_mean + m5_AUC_ci))

# model 6
#m6_bs_perf = m6_bs_perf[is.na(m6_bs_perf$warnings) ,]
m6_bs_perf1 = m6_bs_perf[is.finite(m6_bs_perf$Cutoff) ,]
# average AUC and confidence intervals 
m6_perf <- aggregate(Cutoff~model*Area_under_the_ROC_curve, 
                      mean, data = m6_bs_perf1)
# then calculate mean and CI
m6_AUC_mean <- mean(m6_perf$Area_under_the_ROC_curve)
m6_AUC_sd <- sd(m6_perf$Area_under_the_ROC_curve)
# CI from SD and N
m6_AUC_ci <- v*(m6_AUC_sd/sqrt(N))
m6_AUC_CIs <- cbind((m6_AUC_mean - m6_AUC_ci), (m6_AUC_mean + m6_AUC_ci))

# model 6
#m7_bs_perf = m7_bs_perf[is.na(m7_bs_perf$warnings) ,]
m7_bs_perf1 = m7_bs_perf[is.finite(m7_bs_perf$Cutoff) ,]
# average AUC and confidence intervals 
m7_perf <- aggregate(Cutoff~model*Area_under_the_ROC_curve, 
                      mean, data = m7_bs_perf1)
# then calculate mean and CI
m7_AUC_mean <- mean(m7_perf$Area_under_the_ROC_curve)
m7_AUC_sd <- sd(m7_perf$Area_under_the_ROC_curve)
# CI from SD and N
m7_AUC_ci <- v*(m7_AUC_sd/sqrt(N))
m7_AUC_CIs <- cbind((m7_AUC_mean - m7_AUC_ci), (m7_AUC_mean + m7_AUC_ci))

# get all model AICs 
aic_m1 <- AIC(logLik(mod.m1))
aic_m2 <- AIC(logLik(mod.m2))
aic_m3 <- AIC(logLik(mod.m3))
aic_m4 <- AIC(logLik(mod.m4))
aic_m5 <- AIC(logLik(mod.m5))
aic_m6 <- AIC(logLik(mod.m6))
aic_m7 <- AIC(logLik(mod.m7))
```

# Plot
```{r}
# function to round axes to 1dp
scaleFUN <- function(x) sprintf("%.1f", x)
# simple model
m1ROC_plot <- ggplot() +
  m1_bs_perf %>% 
  geom_line(mapping = aes(x = False_positive_rate, y = True_positive_rate, 
                          group = as.factor(model)),
            alpha = 0.2, colour = "grey85") +
  m1_ROC %>% 
  geom_line(mapping = aes(x = False_positive_rate, y = True_positive_rate),
            colour = dj2[1], size = 1.2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", 
              color = "grey50", size = 1) +
  geom_text(aes(label = paste0("AUC = ",
                               round(m1_ROC[1,4], 2), 
                               ' [', round(m1_AUC_CIs[1], 2), 
                               ', ', round(m1_AUC_CIs[2], 2), ']')), 
            x=.66, y=.02, size = 3.5) +
  geom_text(aes(label = paste0("AIC = ",
                               round(aic_m1, 0))), 
            x=.85, y=.13, size = 3.5) +
  labs(x = 'False positive rate', y = 'True positive rate') +
  scale_y_continuous(labels=scaleFUN) +
  theme_classic() +
  theme(axis.text = element_text(size=axistext),
        axis.title = element_text(size=axistitle),
        title = element_text(size=10)
        )

m4ROC_plot <- ggplot() +
  m4_bs_perf %>% 
  geom_line(mapping = aes(x = False_positive_rate, y = True_positive_rate, 
                          group = as.factor(model)),
            alpha = 0.2, colour = "grey85") +
  m4_ROC %>% 
  geom_line(mapping = aes(x = False_positive_rate, y = True_positive_rate),
            colour = dj2[2], size = 1.2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", 
              color = "grey50", size = 1) +
  geom_text(aes(label = paste0("AUC = ",
                               round(m4_ROC[1,4], 2), 
                               ' [', round(m4_AUC_CIs[1], 2), 
                               ', ', round(m4_AUC_CIs[2], 2), ']')), 
            x=.66, y=.02, size = 3.5) +
  geom_text(aes(label = paste0("AIC = ",
                               round(aic_m4, 0))), 
            x=.85, y=.13, size = 3.5) +
  labs(x = 'False positive rate', y = 'True positive rate') +
  scale_y_continuous(labels=scaleFUN) +
  theme_classic() +
  theme(axis.text = element_text(size=axistext),
        axis.title = element_text(size=axistitle),
        title = element_text(size=10)
        )

m2ROC_plot <- ggplot() +
  m2_bs_perf %>% 
  geom_line(mapping = aes(x = False_positive_rate, y = True_positive_rate, 
                          group = as.factor(model)),
            alpha = 0.2, colour = "grey85") +
  m2_ROC %>% 
  geom_line(mapping = aes(x = False_positive_rate, y = True_positive_rate),
            colour = dj2[3], size = 1.2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", 
              color = "grey50", size = 1) +
  geom_text(aes(label = paste0("AUC = ",
                               round(m2_ROC[1,4], 2), 
                               ' [', round(m2_AUC_CIs[1], 2), 
                               ', ', round(m2_AUC_CIs[2], 2), ']')), 
            x=.66, y=.02, size = 3.5) +
  geom_text(aes(label = paste0("AIC = ",
                               round(aic_m2, 0))), 
            x=.85, y=.13, size = 3.5) +
  scale_y_continuous(labels=scaleFUN) +
  labs(x = 'False positive rate', y = 'True positive rate') +
  theme_classic() +
  theme(axis.text = element_text(size=axistext),
        axis.title = element_text(size=axistitle),
        title = element_text(size=10)
        )

m3ROC_plot <- ggplot() +
  m3_bs_perf %>% 
  geom_line(mapping = aes(x = False_positive_rate, y = True_positive_rate, 
                          group = as.factor(model)),
            alpha = 0.2, colour = "grey85") +
  m3_ROC %>% 
  geom_line(mapping = aes(x = False_positive_rate, y = True_positive_rate),
            colour = dj2[5], size = 1.2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", 
              color = "grey50", size = 1) +
  geom_text(aes(label = paste0("AUC = ",
                               round(m3_ROC[1,4], 2), 
                               ' [', round(m3_AUC_CIs[1], 2), 
                               ', ', round(m3_AUC_CIs[2], 2), ']')), 
            x=.66, y=.02, size = 3.5) +
  geom_text(aes(label = paste0("AIC = ",
                               round(aic_m3, 0))), 
            x=.85, y=.13, size = 3.5) +
  scale_y_continuous(labels=scaleFUN) +
  labs(x = 'False positive rate', y = 'True positive rate') +
  theme_classic() +
  theme(axis.text = element_text(size=axistext),
        axis.title = element_text(size=axistitle),
        title = element_text(size=10)
        )
```

```{r}
# Plot model comparisons
plot_data <- rbind(m1_ROC, m4_ROC, m2_ROC,
                   m3_ROC)
model_order <- c('baseline', 'innocuous TCF \n* noxious TCF', 
                 'innocuous TCF', 'noxious TCF')

plot_data1 <- plot_data %>% 
  arrange(factor(model, levels = model_order))
plot_data1$model <- factor(plot_data$model, levels = model_order)

ROCplot <- ggplot(plot_data1) +
  geom_line(aes(False_positive_rate, True_positive_rate, colour = model), 
                linewidth = 1.5) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", 
              color = "gray50", linewidth = 1) +
  scale_color_manual(values = dj2,
                     labels = model_order) +
  labs(x = '', 
       y = '', 
       title = 'Model comparison') + #labels removed to correspond with below plot
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'none',
        axis.text = element_text(size=axistext),
        axis.title = element_text(size=axistitle),
        title = element_text(size=10)
        )
ROCplot
```

# Marginal effects of all 4 contrast models
```{r}
# plot the effects in the simple and complex (best fitting) model
# re-run model, otherwise cannot get effects
# factor baseline
df_conN$baseline <- as.factor(df_conN$baseline)

# model 1A
eff1 <- ggeffects::ggeffect(mod.m1, 
                           terms = c('baseline'))

eff1_plot <- plot(eff1, ci = FALSE) + 
  geom_point(size = 3, colour = dj2[1]) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                size = 1, width = .3, colour = dj2[1]) +
  labs(title = NULL,
       y = 'Predicted PHS',
       x = 'Contrast condition') +
  #scale_y_continuous(breaks = c(0,.01,.02,.03,.04,.05,.06)) +
  scale_x_continuous(breaks = c(32, 38, 44), labels = c('Low','Medium','High')) +
  theme_classic() +
  theme(axis.text = element_text(size = axistext),
        axis.title = element_text(size = axistitle),
        legend.title = element_text(size = legend),
        legend.text = element_text(size = legend)
        )
eff1_plot
```

# Model 3A (full model)
```{r}
# model 3A
eff2 <- ggeffects::ggeffect(mod.m4, 
                           terms = c('log_TCFdetect','log_TCFpain'))

eff2_plot <- plot(eff2, ci = FALSE) + 
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .1, 
              linewidth = .2) +
  scale_color_manual('Nox. TCF', labels = c('-1SD','mean','+1SD'),
                     values = c(blues9[9],blues9[7],blues9[5])) +
  scale_fill_manual('Nox. TCF', labels = c('-1SD','mean','+1SD'),
                    values = c(blues9[9],blues9[7],blues9[5])) +
  labs(title = NULL,
       y = 'Predicted PHS',
       x = 'Innocuous TCF (log10)') +
  #xlim(-4,0) +
  theme_classic() +
  theme(axis.text = element_text(size = axistext),
        axis.title = element_text(size = axistitle),
        legend.title = element_text(size = legend),
        legend.text = element_text(size = legend),
        legend.position = c(0.30, 0.70),
        strip.text = element_text(size = legend, colour = 'black'))


eff2_plot
```

# Model 3B innocuous only
```{r}
# model 3B
eff3 <- ggeffects::ggeffect(mod.m2, 
                           terms = c('log_TCFdetect'))

eff3_plot <- plot(eff3, ci = FALSE) + 
  geom_line(linewidth = 1.2, colour = dj2[3]) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              colour = dj2[3], fill = dj2[3], alpha = .1, 
              linewidth = .2) +
  labs(title = NULL,
       y = 'Predicted PHS',
       x = 'Innocuous TCF (log10)') +
  #xlim(-4,0) +
  theme_classic() +
  theme(axis.text = element_text(size = axistext),
        axis.title = element_text(size = axistitle),
        )

eff3_plot
                       
```

# Model 3C noxious only
```{r}
# model 3C
eff4 <- ggeffects::ggeffect(mod.m3, terms = c('log_TCFpain'))
    
eff4_plot <- plot(eff4) + 
  geom_line(linewidth = 1.2, colour = dj2[5]) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              colour = dj2[5], fill = dj2[5], alpha = .05, 
              linewidth = .2) +
  labs(title = NULL,
       y = 'Predicted PHS',
       x = 'Noxious TCF (log10)') +
  #xlim(-4,0) +
  theme_classic() +
  theme(axis.text = element_text(size = axistext),
        axis.title = element_text(size = axistitle),
        )

eff4_plot
```
# Combine models and marginal effects
```{r}
alist <- list()
f3a <- ggarrange(m1ROC_plot, NA, eff1_plot,
                 ncol = 3, nrow = 1,
                 widths = c(.8, .05, .6))

f3a <- annotate_figure(f3a, top = text_grob('Model 1A: Contrast condition',
                      size = 12, face = 'bold'))

f3b <- ggarrange(m4ROC_plot, NA, eff2_plot,
                 ncol = 3, nrow = 1,
                 widths = c(.8, .05, .6))
f3b <- annotate_figure(f3b, top = text_grob('Model 3A: Innoc. * nox. TCF',
                      size = 12, face = 'bold'))

f3c <- ggarrange(m2ROC_plot, NA, eff3_plot,
                 ncol = 3, nrow = 1,
                 widths = c(.8, .05, .6))
f3c <- annotate_figure(f3c, top = text_grob('Model 3B: Innocuous TCF',
                      size = 12, face = 'bold'))

f3d <- ggarrange(m3ROC_plot, NA, eff4_plot,
                 ncol = 3, nrow = 1,
                 widths = c(.8, .05, .6))
f3d<- annotate_figure(f3d, top = text_grob('Model 3C: Noxious TCF',
                      size = 12, face = 'bold'))

# combine all

f3 <- ggarrange(f3a, f3b, f3c, f3d,
                ncol = 1, nrow = 4,
                labels = c('A.','B.','C.','D.'))
f3

ggsave('Figure3.png', f3, device = NULL, path = file.path('figures'), 
       width = 5.5, height = 9, dpi = 600)

```

