---
title: "TSL2 Hypothesis 2"
author: "A.G. Mitchell"
date: '2022-07-12'
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE}
# this is getting unweildy, simplify (when you have the time)
knitr::opts_chunk$set(echo = TRUE)

# check for pacman package and install if not found
if (!require("pacman")) install.packages("pacman")
  pacman::p_load(boot, broom.mixed, caret, cluster, factoextra, ggeffects, 
                 gghalves, ggpol, ggpubr, groupdata2, lme4, lmerTest, 
                 RColorBrewer, rcompanion, 
                 reshape2, Rmisc, ROCR, rsample, tidyverse, wesanderson
                 )
```

# Read data from cortex and select data for analysis
```{r data}
# read in data
tsl_file <- file.path("data", "anonymised_tsl2_208.csv")
sum_file <- file.path("data", "summary_tsl2_208.csv")

# load all data file
if (file.exists(tsl_file)){
  tsl2 <- read.csv(tsl_file)
  tsl2 <- tsl2 %>% 
    filter(trials_to_keep == 1)
  } else {
    print('Data file does not exist, check data is in current directory.')
  }
# load data summary file
if (file.exists(sum_file)){
  df <- read.csv(sum_file)
  } else {
    print('Data file does not exist, check data is in current directory if not, run tsl2_phs_q1.Rmd first')
  }
```

# zscore function
```{r}
calc_z <- function (value) {
  z_score <- (value - mean(value))/sd(value)
}
```

# Question 2: pain threshold and PHS count
Investigating whether pain thresholds affect the presence of PHS for both detection and pain conditions. Do this using a mixed effects logistic regression to model the binary outcome of PHS with a random effect of participant, given the individual differences in threshold

# PHS responders
Organise PHS responders from summary data
```{r tsl2}
# extract detection trial only
phs_detect <- df %>% 
  filter(instruction == 'detect') %>% 
  dplyr::rename(phs_detect_bin = phs_01,
         phs_detect = phs)
# extract pain trials only and then add phs_detect to these
phs_pain <- df %>% 
  filter(instruction == 'pain') %>% 
  dplyr::rename(phs_pain_bin = phs_01,
         phs_pain = phs)
# add detection columns to PHS pain
phs_pain1 <- phs_detect %>% 
  left_join(phs_pain, phs_detect, by = c('subject','baseline')) %>%
  select(-c(instruction.y,threshold.y)) %>% 
  dplyr::rename(threshold = threshold.x,
                instruction = instruction.x)

# create large data_frame where only phs during detection is considered
df_detect <- phs_pain1 %>% 
  select(-c(phs_pain, phs_pain_bin))
df_detect <- rbind(df_detect, phs_detect)
```

# Summary statistics - TSL2
```{r sum}
N = 208
# detect by phs
tsl2_mean <- aggregate(threshold~instruction*baseline*phs_01, mean, data = df)
tsl2_median <- aggregate(threshold~instruction*baseline*phs_01, median, data = df)
tsl2_sd <- aggregate(threshold~instruction*baseline*phs_01, sd, data = df)
tsl2_stderr <- tsl2_sd$threshold/sqrt(N)# number of participants
tsl2_ci <- aggregate(threshold~instruction*baseline*phs_01, CI, data = df)
# data frame
tsl2_stats <- data.frame(baseline = tsl2_mean$baseline, phs_detect = tsl2_mean$phs_01, 
                         mean = tsl2_mean$threshold, median = tsl2_median$threshold, 
                         sd = tsl2_sd$threshold, 
                         stderr = tsl2_stderr, tsl2_ci$threshold, 
                         instruction = tsl2_median$instruction)
# rename CI labels
names(tsl2_stats)[names(tsl2_stats) == 'upper'] <- 'ci_upper'
names(tsl2_stats)[names(tsl2_stats) == 'mean.1'] <- 'ci_mean'
names(tsl2_stats)[names(tsl2_stats) == 'lower'] <- 'ci_lower'
# calculate ci
tsl2_stats$ci <- tsl2_stats$ci_upper - tsl2_stats$ci_lower

# all thresholds (regardless of phs)
tsl2_mean <- aggregate(threshold~baseline*instruction, mean, data = df)
tsl2_median <- aggregate(threshold~baseline*instruction, median, data = df)
tsl2_sd <- aggregate(threshold~baseline*instruction, sd, data = df)
tsl2_stderr <- tsl2_sd$threshold/sqrt(N)# number of participants
tsl2_ci <- aggregate(threshold~baseline*instruction, CI, data = df)
# data frame
tsl2_all_stats <- data.frame(baseline = tsl2_mean$baseline, instruction = tsl2_mean$instruction, 
                         mean = tsl2_mean$threshold, median = tsl2_median$threshold, 
                         sd = tsl2_sd$threshold, stderr = tsl2_stderr, tsl2_ci$threshold)
# rename CI labels
names(tsl2_all_stats)[names(tsl2_all_stats) == 'upper'] <- 'ci_upper'
names(tsl2_all_stats)[names(tsl2_all_stats) == 'mean.1'] <- 'ci_mean'
names(tsl2_all_stats)[names(tsl2_all_stats) == 'lower'] <- 'ci_lower'
# calculate ci
tsl2_all_stats$ci <- tsl2_all_stats$ci_upper - tsl2_all_stats$ci_lower
```

# Save all descriptives in results
```{r}
write.csv(tsl2_stats, file.path("results", "phs_threshold_stats.csv"), row.names = FALSE)
tsl2_stats
write.csv(tsl2_all_stats, file.path("results", "all_threshold_stats.csv"), row.names = FALSE)
tsl2_all_stats
```

# Stats for the increase in TSL temperatures with starting temperature
# This does not include PHS
```{r thresholds}
tsl2$trial_z <- calc_z(tsl2$trial)
# innocuous TSL thresholds
m.detect <- lmer(threshold ~ as.factor(baseline) + trial_z + (1|exp_id),
                        data = tsl2 %>% filter(instruction == 'detect'))

summary(m.detect, ddf = "Satterthwaite")

# noxious TSL thresholds
m.pain <- lmer(threshold ~ as.factor(baseline) + trial_z + (1|exp_id),
                      data = tsl2 %>% filter(instruction == 'pain'))

summary(m.pain, ddf = "Satterthwaite")
```

# Threshold logisitic regression
# Prep data
```{r}
# first, correct all pain thresholds < 0 to 0
df2 <- tsl2 %>% 
  dplyr::rename(subject = exp_id)# new data-frame

# create average threshold
avg_threshold <- aggregate(threshold~subject*baseline*instruction, mean, data = df2)
names(avg_threshold)[4] <- 'avg_threshold'

df2 <- merge(df2, avg_threshold, by = c('subject','baseline','instruction'), all.x = TRUE)

# split up pain and detection data sets - is instruction*threshold interaction on phs purely driven by instruction?
df_pain <- df2 %>% 
  filter(instruction == 'pain')
df_detect <- df2 %>% 
  filter(instruction == 'detect')
# change phs to a factor 
df_pain$phs <- as.factor(df_pain$phs)
df_detect$phs <- as.factor(df_detect$phs)

# add pain threshold to detect data
#pain_thresh <- df_pain[, c(1,2,4,7,8)]
pain_thresh <- df_pain %>% 
  select(c(subject,baseline,instruction,threshold,avg_threshold))
df_dp <- merge(df_detect, pain_thresh, by = c('subject', 'baseline'))
# rename thresholds
df_dp<- df_dp %>% 
  dplyr::rename(
    detect_threshold = threshold.x,
    pain_threshold = threshold.y,
    avg_detect_threshold = avg_threshold.x,
    avg_pain_threshold = avg_threshold.y
  )

df_dp$phs <- as.factor(df_dp$phs)
df_dp$baseline <- as.factor(df_dp$baseline)
df_dp$subject <- as.numeric(as.character(df_dp$subject))
# remove unnecessary data
df_dp <- df_dp %>% 
  dplyr::select(c(subject,baseline,gender,trial_z,detect_threshold,phs,avg_detect_threshold,
           pain_threshold,avg_pain_threshold))

# also normalise in original dataset
df_dp <- df_dp %>% 
  mutate(detectz_threshold = calc_z(detect_threshold),
         avg_painz_threshold = calc_z(avg_pain_threshold),
         subject = as.factor(subject))

# density plots
ggplot(df_dp) +
  geom_density(aes(detect_threshold), colour = 'red') +
  geom_density(aes(detectz_threshold))
ggplot(df_dp) +
  geom_density(aes(avg_pain_threshold), colour = 'red') +
  geom_density(aes(avg_painz_threshold)) 

# Create the same data-frame for noxious trials
detect_thresh <- df_detect %>% 
  select(c(subject,baseline,instruction,threshold,avg_threshold))
df_pd <- merge(df_pain, detect_thresh, by = c('subject', 'baseline'))
# rename thresholds
df_pd<- df_pd %>% 
  dplyr::rename(
    detect_threshold = threshold.y,
    pain_threshold = threshold.x,
    avg_detect_threshold = avg_threshold.y,
    avg_pain_threshold = avg_threshold.x
  )

df_pd$phs <- as.factor(df_pd$phs)
df_pd$baseline <- as.factor(df_pd$baseline)
df_pd$subject <- as.numeric(as.character(df_pd$subject))
# remove unnecessary data
df_pd <- df_pd %>% 
   select(c(subject,baseline,gender,trial_z,detect_threshold,phs,avg_detect_threshold,
           pain_threshold,avg_pain_threshold))


# also normalise in original dataset
df_pd <- df_pd %>% 
  mutate(avg_detectz_threshold = calc_z(avg_detect_threshold),
         painz_threshold = calc_z(pain_threshold))

# save both data-sets
write.csv(df_dp, file.path("data", "model_innoc_data.csv"), row.names = FALSE)
write.csv(df_pd, file.path("data", "model_nox_data.csv"), row.names = FALSE)
```

# Then run model :)
```{r}
# more complex model with phs, baseline and thermal thresholds (model of interest)
model.thr <- lme4::glmer(phs~detectz_threshold * avg_painz_threshold + (1|subject),
                 data = df_dp,
                family = 'binomial') #, weights = train_dp$weights)

summary(model.thr)

# calculating odds ratios
# need to find a way of getting CIs from this
or.dp2 <- tidy(model.thr, conf.int=TRUE, 
               exponentiate=TRUE, effects="fixed")
or.dp2
```

# Linear models with added age and gender as predictors
```{r}
# innocuous TSL thresholds
m.detect1 <- lmer(threshold ~ as.factor(baseline) + age + as.factor(gender) + trial_z + (1|exp_id),
                        data = tsl2 %>% filter(instruction == 'detect'))

summary(m.detect1, ddf = "Satterthwaite")

# noxious TSL thresholds
m.pain1 <- lmer(threshold ~ as.factor(baseline) + age + as.factor(gender) + trial_z + (1|exp_id),
                      data = tsl2 %>% filter(instruction == 'pain'))

summary(m.pain1, ddf = "Satterthwaite")
```

