---
title: "TSL2 Supplementary"
author: "A.G. Mitchell"
date: "2022-08-25"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)  # data manipulation
library(rcompanion)
library(reshape2)
library(lme4)
library(lmerTest)
library(gghalves)
library(ggpubr)
library(wesanderson)
library(Rmisc)
library(ROCR)
library(caret)
library(groupdata2)
library(boot)
library(rsample)
library(broom.mixed)
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
library(lavaan) # for SEM
```

# Read data from cortex and select data for analysis
```{r}
Dpath <- '/Users/au706616/Documents/Experiments/PHS-SKULDNET'
filename <- "allsub_tsl2_208.csv" 
tsl2 <- read.csv(file.path(Dpath,filename)) 
```
```{r}
# first, correct all pain thresholds < 0 to 0
N = 208
df2 <- tsl2 %>% 
  filter(task == "tsl2" & trials_to_keep == 1)
df2$threshold[df2$threshold < 0] = 0

# create average threshold
avg_threshold <- aggregate(threshold~exp_id*baseline*instruction, mean, data = df2)
names(avg_threshold)[4] <- 'avg_threshold'

df2 <- merge(df2, avg_threshold, by = c('exp_id','baseline','instruction'), all.x = TRUE)

# split up pain and detection data sets - is instruction*threshold interaction on phs purely driven by instruction?
df_pain <- df2 %>% 
  filter(instruction == 'pain')
df_detect <- df2 %>% 
  filter(instruction == 'detect')
# change phs to a factor 
df_pain$phs <- as.factor(df_pain$phs)
df_detect$phs <- as.factor(df_detect$phs)
```

# SECOND MODEL COMPARISON: Pain and detection thresholds on noxious phs
Run the same as above but this time to predict the likelihood of noxious phs

```{r}
# add pain threshold to detect data
#detect_thresh <- df_detect[, c(1,2,4,7,8)]
# add pain threshold to detect data
#pain_thresh <- df_pain[, c(1,2,4,7,8)]
detect_thresh <- df_detect[, c(1,2,10,11,16)]
df_pd <- merge(df_pain, detect_thresh, by = c('exp_id', 'trial', 'baseline'))
# rename thresholds
df_pd <- df_pd %>% 
  dplyr::rename(
    detect_threshold = threshold.y,
    pain_threshold = threshold.x,
    avg_detect_threshold = avg_threshold.y,
    avg_pain_threshold = avg_threshold.x
  )

df_pd$phs <- as.factor(df_pd$phs)
df_pd$baseline <- as.factor(df_pd$baseline)
df_pd$exp_id <- as.numeric(as.character(df_pd$exp_id))
# remove unnecessary data
df_pd <- df_pd[, c(1:4,11,14,17,18)]

# also normalise in original dataset
df_pd$detectz_threshold <- (df_pd$detect_threshold - 
                                 mean(df_pd$detect_threshold))/sd(df_pd$detect_threshold)
df_pd$painz_threshold <- (df_dp$pain_threshold - 
                                mean(df_dp$pain_threshold))/sd(df_pd$pain_threshold)
df_pd$avg_detectz_threshold <- (df_pd$avg_detect_threshold - 
                                mean(df_pd$avg_detect_threshold))/sd(df_pd$avg_detect_threshold)

# density plots
ggplot(df_pd) +
  geom_density(aes(pain_threshold), colour = 'red') +
  geom_density(aes(painz_threshold))
ggplot(df_pd) +
  geom_density(aes(avg_detect_threshold), colour = 'red') +
  geom_density(aes(avg_detectz_threshold)) 
```

# First run the simple model, noxious PHS and baseline temp
# Run the model and prediction
```{r}
# run logistic regression on detection phs baseline
model.pd1 <- lme4::glmer(phs~as.ordered(baseline) + (1|exp_id),
                 data = df_pd,
                    family = 'binomial')
summary(model.pd1)

# calculating odds ratios of fixed effects
or.pd1 <- tidy(model.pd1, conf.int=TRUE, exponentiate=TRUE, effects="fixed")
# convert the odds ratio to a percentage
or.pd1$percent <- or.pd1$estimate/(1-or.pd1$estimate)
```

# Then second, more complex model with threshold added
```{r}
# run logistic regression on noxious PHs, baseline and thresholds (complex)
model.pd2 <- lme4::glmer(phs~as.ordered(baseline) 
                         + painz_threshold + avg_detectz_threshold + (1|exp_id),
                 data = df_pd,
                 family = 'binomial')
summary(model.pd2)

# calculating odds ratios of fixed effects
or.pd2 <- tidy(model.pd2, conf.int=TRUE, exponentiate=TRUE, effects="fixed")
# convert the odds ratio to a percentage
or.pd2$percent <- or.pd2$estimate/(1-or.pd2$estimate)
```

*Model comparisons*
Functions to extract data from objects of class performance  created with ROCR
# Conduct sensitivity and ROC analyses on the simple model
```{r}
#load get_ROCR_data function from file
source("get_ROCR_data.R")

# predictions with a simple model
hist(predict(model.pd1, re.form = ~0))
pred_resp1 <- predict(model.pd1, type="response", re.form = ~0) #re.form = ~0 do not include random effects, default is NULL which includes all re
hist(pred_resp1)

# Create performance object for the simple model and extract the data.
model.pd1_pred_obj <- prediction(predictions = pred_resp1, 
                                 labels = df_pd$phs)
model.pd1_perf_roc <- performance(prediction.obj = model.pd1_pred_obj, 
                                  measure = "tpr",
                                  x.measure = "fpr")
model.pd1_perf_auc <- performance(prediction.obj = model.pd1_pred_obj, 
                                  measure = "auc")
model.pd1_perf <- cbind(get_ROCR_data(model.pd1_perf_roc),
                        get_ROCR_data(model.pd1_perf_auc))

model.pd1_perf$model <- "baseline" #add model identification for plotting
```

# Sensitivity analysis for complex model
```{r}
# then again for the complex model
hist(predict(model.pd2, re.form = ~0))
pred_resp2 <- predict(model.pd2, type="response", re.form = ~0)
hist(pred_resp2)

model.pd2_pred_obj <- prediction(predictions = pred_resp2, 
                                 labels = df_pd$phs)
model.pd2_perf_roc <- performance(prediction.obj = model.pd2_pred_obj, 
                                  measure = "tpr",
                                  x.measure = "fpr")
model.pd2_perf_auc <- performance(prediction.obj = model.pd2_pred_obj, 
                                  measure = "auc")
model.pd2_perf <- cbind(get_ROCR_data(model.pd2_perf_roc),
                        get_ROCR_data(model.pd2_perf_auc))

model.pd2_perf$model <- "baseline \n+ detectz_threshold \n+ avg_painz_threshold" #add model identification for plotting
```

# Plot the comparison
```{r}
plot_data <- rbind(model.pd1_perf, model.pd2_perf)

ROCplot <- ggplot(plot_data) +
  geom_line(aes(False_positive_rate, True_positive_rate, colour = model), 
                size = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", 
              color = "gray50", size = .75) +
  scale_color_manual(values = c('grey70', 'black'),
                     labels = c('Pre-warming only', 'Pre-warming, detection & pain thresholds')) +
  labs(x = 'False positive rate', 
       y = 'True positive rate', 
       title = 'Noxious PHS ROC model comparison') +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        axis.text = element_text(size = 10))
ROCplot

ggsave('NOX_model-comparison.tiff', ROCplot, path = Dpath, 
       width = 6, height = 5, dpi = 600)
```

# Bootstrap both model fits (simple and complex) to get an idea of confidence of AUC
Use 20 replicates during code testing, 2000 for data analysis and model checks
Use the same function as in Chunk 10

```{r}
#identify ids for bootstrap replicates
B <- 2000 #number of bootstrap replicates
ids <- unique(df_pd$exp_id) #unique ids
N <- length(ids) #sample size

#create a list of per subject mini datasets to use for bs
id_sets_list <- list() #list to store mini-datasets,m separete one for each subject, empty atm
#populate id_sets_list
for (id in ids) {
  id_sets_list[[id]] <- df_pd[df_pd$exp_id == id, ]
}

#create datasets for bootstrapping using rsample package
bs_train <- list() #empty list to store datasets
bs_test <- list() 
set.seed(493) #so it is reproducible
for (i in seq.int(B)) {
  bs_train[[i]] <- bind_rows(sample(id_sets_list, N, replace = TRUE))
  bs_test[[i]] <- bind_rows(sample(id_sets_list, N, replace = TRUE))
}
```

# Function to get a dataset similar to df_dp, fits a model and extracts data 
# from ROCR performance objects
# Should also work on a df list 

```{r}
#extracted data: tpr, fpr, auc
get_performance_df <- function (train, test) {
  # calculate zscores
  train$detectz_threshold <- (train$avg_detect_threshold - 
                                 mean(train$avg_detect_threshold))/sd(train$avg_detect_threshold)
  train$painz_threshold <- (train$pain_threshold - 
                                 mean(train$pain_threshold))/sd(train$pain_threshold)
  # renormalise data - testing
  test$detectz_threshold <- (test$avg_detect_threshold - 
                                 mean(test$avg_detect_threshold))/sd(test$avg_detect_threshold)
  test$painz_threshold <- (test$pain_threshold - 
                                 mean(test$pain_threshold))/sd(test$pain_threshold)
 
  # run models
  # complex including thresholds
  complex.model <- lme4::glmer(phs ~ as.ordered(baseline) + detectz_threshold + 
                         painz_threshold + (1|exp_id),
                       data = train,
                       family = 'binomial')
   
  # make sure warnings get logged - remove simulation later
  warning <- as.character(summary(complex.model)$optinfo$conv$lme4$messages)
  if(length(warning) != 0) print(warning)
  warning <- ifelse(length(warning) == 0, NA, warning)
  
  # predictions from the complex model
  preds <- predict(complex.model, test, type="response", re.form = ~0)
  preds_obj <- prediction(preds, test$phs)
  roc_obj <- performance(preds_obj, measure = "tpr", x.measure = "fpr")
  auc_obj <- performance(preds_obj, measure = "auc")
  com_performance_df <- cbind(get_ROCR_data(roc_obj),
                          get_ROCR_data(auc_obj))
  com_performance_df$warnings <- warning
  
  # simple model
  simple.model <- lme4::glmer(phs ~ as.ordered(baseline) + (1|exp_id),
                       data = train,
                       family = 'binomial')
  warning <- as.character(summary(simple.model)$optinfo$conv$lme4$messages)
  if(length(warning) != 0) print(warning)
  warning <- ifelse(length(warning) == 0, NA, warning)
  
  # predictions from the complex model
  preds <- predict(simple.model, test, type="response", re.form = ~0)
  preds_obj <- prediction(preds, test$phs)
  roc_obj <- performance(preds_obj, measure = "tpr", x.measure = "fpr")
  auc_obj <- performance(preds_obj, measure = "auc")
  sim_performance_df <- cbind(get_ROCR_data(roc_obj),
                          get_ROCR_data(auc_obj))
  sim_performance_df$warnings <- warning
  
  performance_df <- list(sim_performance_df, com_performance_df)
  
  return(performance_df)
}
```

# Run bootstrapping to get idea of model performance
```{r, warning = FALSE, message = FALSE}
# now get a list of dfs and get_performance_df function and get bs results!
# do this for both models
com_bs_performance <- data.frame(True_positive_rate = numeric(), 
                             False_positive_rate = numeric(), 
                             Cutoff = numeric(), 
                             AUC = numeric(), 
                             model = numeric(),
                             warnings = character())
sim_bs_performance <- data.frame(True_positive_rate = numeric(), 
                             False_positive_rate = numeric(), 
                             Cutoff = numeric(), 
                             AUC = numeric(), 
                             model = numeric(),
                             warnings = character())
# running bootstrapping on both simple and complex models
for (i in seq.int(B)) {
  performance <- get_performance_df(bs_train[[i]], bs_test[[i]])
  performance[[1]]$model <- i
  performance[[2]]$model <- i
  # extract data from simulation
  sim_bs_performance <- rbind(sim_bs_performance, performance[[1]])
  com_bs_performance <- rbind(com_bs_performance, performance[[2]])
}

# save data-frames specific to innocuous analysis
nox_sim_bs_perf <- sim_bs_performance
nox_com_bs_perf <- com_bs_performance
```

# Plotting the results of the bootstrap and obtaining AUC
```{r}
# first remove models with warnings - both data-frames
nox_sim_bs_perf = nox_sim_bs_perf[is.na(nox_sim_bs_perf$warnings) ,]
sim_bs_perf = nox_sim_bs_perf[is.finite(nox_sim_bs_perf$Cutoff) ,] # get rid of infinite values

nox_com_bs_perf = nox_com_bs_perf[is.na(nox_com_bs_perf$warnings) ,]
com_bs_perf = nox_com_bs_perf[is.finite(nox_com_bs_perf$Cutoff) ,] # get rid of infinite values

# average AUC and confidence intervals - simple model
sim_perf <- aggregate(Cutoff~model*Area_under_the_ROC_curve, 
                      mean, data = sim_bs_perf)
# then calculate mean and CI
sim_AUC_mean <- mean(sim_perf$Area_under_the_ROC_curve)
sim_AUC_CI <- CI(sim_perf$Area_under_the_ROC_curve)

# average AUC and confidence intervals - complex model
com_perf <- aggregate(Cutoff~model*Area_under_the_ROC_curve, 
                      mean, data = com_bs_perf)
# then calculate mean and CI
com_AUC_mean <- mean(com_perf$Area_under_the_ROC_curve)
com_AUC_CI <- CI(com_perf$Area_under_the_ROC_curve)
```

# Plotting the AUC and CIs
```{r}
# simple model
simROC_plot <- ggplot() +
  nox_sim_bs_perf %>% 
  geom_line(mapping = aes(x = False_positive_rate, y = True_positive_rate, 
                          group = as.factor(model)),
            alpha = 0.2, colour = "grey50") +
  model.pd1_perf %>% 
  geom_line(mapping = aes(x = False_positive_rate, y = True_positive_rate),
            colour = "grey10", size = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", color = "grey") +
  geom_text(aes(label = paste0('AUC MOD = ', round(model.pd1_perf[1,4], 2))),
            x=.75, y=.075, size = 4) +
  geom_text(aes(label = paste0('AUC SIM = ', round(sim_AUC_mean, 2), 
                               ' [', round(sim_AUC_CI[1], 2), 
                               ', ', round(sim_AUC_CI[3], 2), ']')), 
            x=.825, y=.01, size = 4) +
  labs(title = 'Model: phs ~ pre-warming', 
       x = '', y = 'True positive rate') +
  theme_classic() +
  theme(axis.text = element_text(size = 10))

# complex model
comROC_plot <- ggplot() +
  nox_com_bs_perf %>% 
  geom_line(mapping = aes(x = False_positive_rate, y = True_positive_rate, group = as.factor(model)),
            alpha = 0.2, colour = "grey50") +
  model.pd2_perf %>% 
  geom_line(mapping = aes(x = False_positive_rate, y = True_positive_rate),
            colour = "grey10", size = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", color = "grey") +
  geom_text(aes(label = paste0('AUC MOD = ', round(model.pd2_perf[1,4], 2),
                               '0')),
            x=.75, y=.075, size = 4) +
  geom_text(aes(label = paste0('AUC SIM = ', round(com_AUC_mean, 2), 
                               ' [', round(com_AUC_CI[1], 2), 
                               ', ', round(com_AUC_CI[3], 2), ']')), 
            x=.825, y=.01, size = 4) +
  labs(title = 'Model: phs ~ pre-warming + detection threshold + pain threhsold', 
       x = 'False positive rate', y = 'True positive rate') +
  theme_classic() +
  theme(axis.text = element_text(size = 10))

NOX_ROCplot_CI <- ggarrange(simROC_plot, comROC_plot,
                        ncol = 1, nrow = 2,
                        labels = c('A.', 'B.'))
NOX_ROCplot_CI

ggsave('NOX_ROCplots_CI.tiff', ROCplot_CI, device = NULL, path = Dpath, 
       width = 6, height = 8, dpi = 600)
```

# Plotting - noxious phs
```{r}
# Plot phs pain, detection and pain thresholds
## Plotting - raincloud
# preparing average data-set
df_pd1 <- aggregate(detect_threshold~baseline*exp_id, mean, data = df_pd) # avg detection thresh
df_pd2 <- aggregate(pain_threshold~baseline*exp_id, mean, data = df_pd) # avg pain thresh
df_pd$phs_01 <- as.numeric(as.character(df_pd$phs)) #as numeric
df_pd3 <- aggregate(phs_01~baseline*exp_id, max, data = df_pd) # avg phs yes/no
df_pd4 <- aggregate(phs_01~baseline*exp_id, sum, data = df_pd) # avg phs count
names(df_pd4)[3] <- 'phs_all' # renaming
# merge all
df_nox <- merge(df_pd1, df_pd2, by = c('exp_id','baseline'))
df_nox <- merge(df_nox, df_pd3, by = c('exp_id','baseline'))
df_nox <- merge(df_nox, df_pd4, by = c('exp_id','baseline'))
df_nox$baseline <- as.numeric(df_nox$baseline)

# summary statistics 
# summary statistics
detect_nox_mean <- summarySEwithin(data = df_nox, measurevar = 'detect_threshold', 
                               withinvars = c('baseline','phs_01'))
pain_nox_mean <- summarySEwithin(data = df_nox, measurevar = 'pain_threshold', 
                               withinvars = c('baseline','phs_01'))
detect_nox_mean$baseline <- as.numeric(detect_nox_mean$baseline)
pain_nox_mean$baseline <- as.numeric(pain_nox_mean$baseline)
```

# Plotting figure 3A - noxious PHS and detection thresholds
```{r}
library(RColorBrewer)
gren <- brewer.pal(6, "Greens")
blue <- brewer.pal(6, "Blues")
reds <- brewer.pal(6, "Reds")
purp <- brewer.pal(6, "Purples")
grey <- brewer.pal(6, "Greys")
oran <- brewer.pal(6, "Oranges")

nudge1 = -0.1
nudge2 = 0.1

# jitter
df_nox$xj <- jitter(df_nox$baseline, amount = .1)

# plot phs detection thresolds
f3a <- ggplot(data = df_nox, 
             mapping = aes(x = as.factor(baseline), y = detect_threshold, 
                           group = phs_01, fill = as.factor(phs_01), 
                           colour = as.factor(phs_01))) +
  #geom_point(aes(x = xj), shape = 21, size = 2, alpha = .6) +
  # Add rainclouds - no PHS
  geom_half_violin(
      data = df_nox %>% 
      filter(phs_01 == 0, baseline == 1), 
      aes(x = baseline, y = detect_threshold), 
      position = position_nudge(x = nudge1), side = "l", alpha = .6
      ) +
  geom_point(
    data = detect_nox_mean %>% 
      filter(phs_01 == 0, baseline == 1),
      aes(x = baseline, y = detect_threshold),
      position = position_nudge(x = nudge1+.05), size = 2) +
  geom_errorbar(
    data = detect_nox_mean %>% 
      filter(phs_01 == 0, baseline == 1),
    aes(x = baseline, ymin = detect_threshold-ci, ymax = detect_threshold+ci),
    width = .05, position = position_nudge(x = nudge1+.05), size = .5
  ) +
  geom_half_violin(
      data = df_nox %>% 
      filter(phs_01 == 0, baseline == 2), 
      aes(x = baseline, y = detect_threshold),
      position = position_nudge(x = nudge1), side = "l", alpha = .6
      ) +
  geom_point(
    data = detect_nox_mean %>% 
      filter(phs_01 == 0, baseline == 2),
      aes(x = baseline, y = detect_threshold),
      position = position_nudge(x = nudge1+.05), size = 2) +
  geom_errorbar(
    data = detect_nox_mean %>% 
      filter(phs_01 == 0, baseline == 2),
    aes(x = baseline, ymin = detect_threshold-ci, ymax = detect_threshold+ci),
    width = .05, position = position_nudge(x = nudge1+.05), size = .5
  ) +
  geom_half_violin(
      data = df_nox %>% 
      filter(phs_01 == 0, baseline == 3), 
      aes(x = baseline, y = detect_threshold),
      position = position_nudge(x = nudge1), side = "l", alpha = .6
      ) +
  geom_point(
    data = detect_nox_mean %>% 
      filter(phs_01 == 0, baseline == 3),
      aes(x = baseline, y = detect_threshold),
      position = position_nudge(x = nudge1+.05), size = 2) +
  geom_errorbar(
    data = detect_nox_mean %>% 
      filter(phs_01 == 0, baseline == 3),
    aes(x = baseline, ymin = detect_threshold-ci, ymax = detect_threshold+ci),
    width = .05, position = position_nudge(x = nudge1+.05), size = .5
  ) +
  # PHS
  geom_half_violin(
      data = df_nox %>% 
      filter(phs_01 == 1, baseline == 1), 
      aes(x = baseline, y = detect_threshold), 
      position = position_nudge(x = nudge2), side = "r", alpha = .6
      ) +
  geom_point(
    data = detect_nox_mean %>% 
      filter(phs_01 == 1, baseline == 1),
      aes(x = baseline, y = detect_threshold),
      position = position_nudge(x = nudge2-.05), size = 2) +
  geom_errorbar(
    data = detect_nox_mean %>% 
      filter(phs_01 == 1, baseline == 1),
    aes(x = baseline, ymin = detect_threshold-ci, ymax = detect_threshold+ci),
    width = .05, position = position_nudge(x = nudge2-.05), size = .5
  ) +
  geom_half_violin(
      data = df_nox %>% 
      filter(phs_01 == 1, baseline == 2), 
      aes(x = baseline, y = detect_threshold),
      position = position_nudge(x = nudge2), side = "r", alpha = .6
      ) +
  geom_point(
    data = detect_nox_mean %>% 
      filter(phs_01 == 1, baseline == 2),
      aes(x = baseline, y = detect_threshold),
      position = position_nudge(x = nudge2-.05), size = 2) +
  geom_errorbar(
    data = detect_nox_mean %>% 
      filter(phs_01 == 1, baseline == 2),
    aes(x = baseline, ymin = detect_threshold-ci, ymax = detect_threshold+ci),
    width = .05, position = position_nudge(x = nudge2-.05), size = .5
  ) +
  geom_half_violin(
      data = df_nox %>% 
      filter(phs_01 == 1, baseline == 3), 
      aes(x = baseline, y = detect_threshold),
      position = position_nudge(x = nudge2), side = "r", alpha = .6
      ) +
  geom_point(
    data = detect_nox_mean %>% 
      filter(phs_01 == 1, baseline == 3),
      aes(x = baseline, y = detect_threshold),
      position = position_nudge(x = nudge2-.05), size = 2) +
   geom_errorbar(
    data = detect_nox_mean %>% 
      filter(phs_01 == 1, baseline == 3),
    aes(x = baseline, ymin = detect_threshold-ci, ymax = detect_threshold+ci),
    width = .05, position = position_nudge(x = nudge2-.05), size = .5
  ) +
  scale_fill_manual(labels = c("No PHS", "PHS"), values=c(grey[1],oran[2])) +
  scale_colour_manual(labels = c("No PHS", "PHS"), values=c(grey[5],oran[6])) +
  # Define additional settings
  scale_x_continuous(breaks=c(1,2,3), labels=c("32", "38", "44"), ) +
  labs(title = NULL, 
       y = "Detection threshold (ºC)",
       x = NULL) +
  #ylim(15,35) +
  theme_classic() +
  theme(legend.position = 'none')
f3a
```

# Plot figure 3B - Noxious PHS and pain thresholds
```{r}
# plot phs pain thresholds
f3b <- ggplot(data = df_nox, 
             mapping = aes(x = as.factor(baseline), y = pain_threshold, 
                           group = phs_01, fill = as.factor(phs_01), 
                           colour = as.factor(phs_01))) +
  #geom_point(aes(x = xj), shape = 21, size = 2, alpha = .6) +
  # Add rainclouds - no PHS
  geom_half_violin(
      data = df_nox %>% 
      filter(phs_01 == 0, baseline == 1), 
      aes(x = baseline, y = pain_threshold), 
      position = position_nudge(x = nudge1), side = "l", alpha = .6
      ) +
  geom_point(
    data = pain_nox_mean %>% 
      filter(phs_01 == 0, baseline == 1),
      aes(x = baseline, y = pain_threshold),
      position = position_nudge(x = nudge1+.05), size = 2) +
  geom_errorbar(
    data = pain_nox_mean %>% 
      filter(phs_01 == 0, baseline == 1),
    aes(x = baseline, ymin = pain_threshold-ci, ymax = pain_threshold+ci),
    width = .05, position = position_nudge(x = nudge1+.05), size = .5
  ) +
  geom_half_violin(
      data = df_nox %>% 
      filter(phs_01 == 0, baseline == 2), 
      aes(x = baseline, y = pain_threshold),
      position = position_nudge(x = nudge1), side = "l", alpha = .6
      ) +
  geom_point(
    data = pain_nox_mean %>% 
      filter(phs_01 == 0, baseline == 2),
      aes(x = baseline, y = pain_threshold),
      position = position_nudge(x = nudge1+.05), size = 2) +
  geom_errorbar(
    data = pain_nox_mean %>% 
      filter(phs_01 == 0, baseline == 2),
    aes(x = baseline, ymin = pain_threshold-ci, ymax = pain_threshold+ci),
    width = .05, position = position_nudge(x = nudge1+.05), size = .5
  ) +
  geom_half_violin(
      data = df_nox %>% 
      filter(phs_01 == 0, baseline == 3), 
      aes(x = baseline, y = pain_threshold),
      position = position_nudge(x = nudge1), side = "l", alpha = .6
      ) +
  geom_point(
    data = pain_nox_mean %>% 
      filter(phs_01 == 0, baseline == 3),
      aes(x = baseline, y = pain_threshold),
      position = position_nudge(x = nudge1+.05), size = 2) +
  geom_errorbar(
    data = pain_nox_mean %>% 
      filter(phs_01 == 0, baseline == 3),
    aes(x = baseline, ymin = pain_threshold-ci, ymax = pain_threshold+ci),
    width = .05, position = position_nudge(x = nudge1+.05), size = .5
  ) +
  # PHS
  geom_half_violin(
      data = df_nox %>% 
      filter(phs_01 == 1, baseline == 1), 
      aes(x = baseline, y = pain_threshold), 
      position = position_nudge(x = nudge2), side = "r", alpha = .6
      ) +
  geom_point(
    data = pain_nox_mean %>% 
      filter(phs_01 == 1, baseline == 1),
      aes(x = baseline, y = pain_threshold),
      position = position_nudge(x = nudge2-.05), size = 2) +
  geom_errorbar(
    data = pain_nox_mean %>% 
      filter(phs_01 == 1, baseline == 1),
    aes(x = baseline, ymin = pain_threshold-ci, ymax = pain_threshold+ci),
    width = .05, position = position_nudge(x = nudge2-.05), size = .5
  ) +
  geom_half_violin(
      data = df_nox %>% 
      filter(phs_01 == 1, baseline == 2), 
      aes(x = baseline, y = pain_threshold),
      position = position_nudge(x = nudge2), side = "r", alpha = .6
      ) +
  geom_point(
    data = pain_nox_mean %>% 
      filter(phs_01 == 1, baseline == 2),
      aes(x = baseline, y = pain_threshold),
      position = position_nudge(x = nudge2-.05), size = 2) +
  geom_errorbar(
    data = pain_nox_mean %>% 
      filter(phs_01 == 1, baseline == 2),
    aes(x = baseline, ymin = pain_threshold-ci, ymax = pain_threshold+ci),
    width = .05, position = position_nudge(x = nudge2-.05), size = .5
  ) +
  geom_half_violin(
      data = df_nox %>% 
      filter(phs_01 == 1, baseline == 3), 
      aes(x = baseline, y = pain_threshold),
      position = position_nudge(x = nudge2), side = "r", alpha = .6
      ) +
  geom_point(
    data = pain_nox_mean %>% 
      filter(phs_01 == 1, baseline == 3),
      aes(x = baseline, y = pain_threshold),
      position = position_nudge(x = nudge2-.05), size = 2) +
   geom_errorbar(
    data = pain_nox_mean %>% 
      filter(phs_01 == 1, baseline == 3),
    aes(x = baseline, ymin = pain_threshold-ci, ymax = pain_threshold+ci),
    width = .05, position = position_nudge(x = nudge2-.05), size = .5
  ) +
  scale_fill_manual(labels = c("No PHS", "PHS"), values=c(grey[1],oran[2])) +
  scale_colour_manual(labels = c("No PHS", "PHS"), values=c(grey[5],oran[6])) +
  # Define additional settings
  scale_x_continuous(breaks=c(1,2,3), labels=c("32", "38", "44"), ) +
  labs(title = NULL, 
       y = "Pain threshold (ºC)",
       x = "Pre-warming temperature (ºC)") +
  #ylim(15,35) +
  theme_classic() +
  theme(legend.position = 'none')
f3b
```

```{r}
f3 <- ggarrange(f3a, f3b,
                ncol = 1, nrow = 2,
                labels = c('A', 'B'))
f3

ggsave('Figure3.tiff', f3, path = Dpath, 
       width = 6.5, height = 7, dpi = 700)
  
```

# Plot noxious summary statstics with baseline
```{r}
# summary statistics
sum_pd_detect <- summarySEwithin(data = df_nox, measurevar = 'detect_threshold', 
                          withinvars = c('phs_01', 'baseline'))
# rename
sum_pd_detect <- sum_pd_detect %>% 
  dplyr::rename(detect_sd = sd,
         detect_se = se,
         detect_ci = ci)
sum_pd_pain <- summarySEwithin(data = df_nox, measurevar = 'pain_threshold', 
                          withinvars = c('phs_01', 'baseline'))
# rename
sum_pd_pain <- sum_pd_pain %>% 
  dplyr::rename(pain_sd = sd,
         pain_se = se,
         pain_ci = ci)
# merge the two
sum_pd <- merge(sum_pd_detect, sum_pd_pain, by = c('phs_01','baseline','N'))

# plotting
# detection
f5a <- ggplot(sum_pd, aes(group = phs_01, colour = phs_01)) +
  geom_point(aes(baseline, detect_threshold), size = 2.5,
             position = position_dodge(.2)) +
  geom_line(aes(baseline, detect_threshold),
            position = position_dodge(.2), size = .8) +
  geom_errorbar(aes(baseline, ymin=detect_threshold-detect_ci, 
                    ymax=detect_threshold+detect_ci),
                position = position_dodge(.2), width = .15, size = .8) +
  labs(y = 'Detection threshold (ºC)', x = 'Baseline temperature (ºC)') +
  ylim(25,32) +
  scale_colour_manual(labels = c("No PHS", "PHS"), values = c(purp[3], purp[5]))+
  theme_classic() +
  theme(legend.position = 'bottom',
        legend.title = element_blank())

# pain
f5b <- ggplot(sum_pd, aes(group = phs_01, colour = phs_01)) +
  geom_point(aes(baseline, pain_threshold), size = 2.5,
             position = position_dodge(.2)) +
  geom_line(aes(baseline, pain_threshold),
            position = position_dodge(.2), size = .8) +
  geom_errorbar(aes(baseline, ymin=pain_threshold-pain_ci, 
                    ymax=pain_threshold+pain_ci),
                position = position_dodge(.2), width = .15, size = .8) +
  labs(y = 'Pain threshold (ºC)', x = 'Baseline temperature (ºC)') +
  ylim(8,24) +
  scale_colour_manual(labels = c("No PHS", "PHS"), values = c(reds[3], reds[5]))+
  theme_classic() +
  theme(legend.position = 'none',
        legend.title = element_blank())

f5 <- ggarrange(f5a, f5b,
                ncol = 2, nrow = 1,
                common.legend = TRUE)

f5

ggsave('NoxPHS_thresh-baseline.tiff', f5, width = 6, height = 4, dpi = 300)

```

# Latent curve modelling for innocuous cold range - NOXIOUS PHS
Now do the same for noxious compared to innocious PHS

```{r}
# Prepare the data
# select specific trials for analysis
df3 <- df

# calculating the innocuous cold range for each trial
df_detect2 <- df3 %>% 
  filter(instruction == 'detect')
# then aggregate across trials, so not using trial-by-trial data
df_detect3 <- aggregate(threshold~exp_id+baseline, mean, data = df_detect2)
df_detect4 <- aggregate(phs~exp_id+baseline, sum, data = df_detect2)
# merge
df_detect2 <- merge(df_detect3, df_detect4, by = c('exp_id', 'baseline'))

df_pain2 <- df3 %>% 
  filter(instruction == 'pain')
# then aggregate across trials, so not using trial-by-trial data
df_pain3 <- aggregate(threshold~exp_id+baseline, mean, data = df_pain2)
df_pain4 <- aggregate(phs~exp_id+baseline, sum, data = df_pain2)
# merge
df_pain2 <- merge(df_pain3, df_pain4, by = c('exp_id', 'baseline'))

diff2 <- df_detect2$threshold - df_pain2$threshold

# getting data
df3 <- data.frame(subject = df_pain2$exp_id, baseline = df_pain2$baseline, 
                  phs = df_pain2$phs, measure = rep("detect-pain", nrow(df_pain2)), 
                  difference = diff2)
# create a binary phs
df3$phs_01 <- ifelse(df3$phs > 0, 1, 0)

# grouping by baseline- take the standardised values
df_32 <- df3 %>% 
  filter(baseline == 32) %>% 
  dplyr::rename(icr32 = difference,
                phs32 = phs,
                phs32_01 = phs_01)
df_38 <- df3 %>% 
  filter(baseline == 38) %>% 
  dplyr::rename(icr38 = difference,
                phs38 = phs,
                phs38_01 = phs_01)
df_44 <- df3 %>% 
  filter(baseline == 44) %>% 
  dplyr::rename(icr44 = difference,
                phs44 = phs,
                phs44_01 = phs_01)
# merge
df_lcm_p <- merge(df_32, df_38, by = c('subject','measure'))
df_lcm_p <- merge(df_lcm_p, df_44, by = c('subject','measure'))
# select relevant columns
df_lcm_p <- df_lcm_p[, c(1,2,4:6,8:10,12:14)]

# as factors
df_lcm_p$phs32 <- as.factor(df_lcm_p$phs32)
df_lcm_p$phs38 <- as.factor(df_lcm_p$phs38)
df_lcm_p$phs44 <- as.factor(df_lcm_p$phs44)
df_lcm_p$phs32_01 <- as.factor(df_lcm_p$phs32_01)
df_lcm_p$phs38_01 <- as.factor(df_lcm_p$phs38_01)
df_lcm_p$phs44_01 <- as.factor(df_lcm_p$phs44_01)
```

# Latent growth curve model of ICR with PHS added as time variant predictor variable

```{r}
# different model
phs_var.model3 <- '
i =~ 1*icr32 + 1*icr38 + 1*icr44
s =~ 0*icr32 + 1*icr38 + 2*icr44

icr32 ~ phs32_01
icr38 ~ phs38_01
icr44 ~ phs44_01

'
# run model
phs3.fit <- growth(phs_var.model3, data = df_lcm_p, estimator = 'MLR')
summary(phs3.fit, fit.measures = TRUE, standardized = TRUE)
```

# contrast pain mods
# Pain models
```{r}
# same for pain
df_conP <- df_con %>% filter(instruction == 'pain')
# rescale variables first
df_conP$z_contrast <- (df_conP$contrast - mean(df_conP$contrast))/sd(df_conP$contrast)

# plot rescaled
ggplot(data = df_conP) +
  geom_density(aes(z_contrast)) +
  facet_wrap(~baseline) +
  labs(title = 'Distribution of noxious contrast by starting temperature',
       x = 'Noxious Thermal Contrast (z-score)',
       y = 'Density') +
  theme_bw()

# run model
model.conP <- lme4::glmer(phs ~ baseline * z_contrast + trial + (1|subject),
                data = df_conP,
                family = 'binomial') #, weights = train_dp$weights)

summary(model.conP)

# calculating odds ratios
or.conP <- tidy(model.conP, conf.int=TRUE, 
               exponentiate=TRUE, effects="fixed")
or.conP
```

# Get one contrast value per participant for detect and pain
```{r}
# summarise phs
phs_all <- aggregate(phs~subject*instruction, sum, data = phs_count)
phs_01 <- as.numeric(phs_all$phs != 0) #binarise
# merge
phs_all <- cbind(phs_all, phs_01)

# then combine with average contrasts
con_pp <- merge(con_pp, phs_all, by = c('subject','instruction'))

# summary statistics
con_sum <- summarySE(data = con_pp, measurevar = 'contrast', 
                     groupvars = c('instruction','phs_01'), conf.interval = 0.95)

# plot :)
con_pp$xj <- jitter(con_pp$phs_01, amount = .05)
# innocuous
ggplot() +
  geom_point(data = con_pp %>% filter(instruction == 'detect'), 
       aes(xj, contrast), alpha = .5, size = 2) +
  geom_half_violin(data = con_pp %>% 
                     filter(instruction == 'detect' & phs_01 == 0),
                   aes(phs_01, contrast),
                   position = position_nudge(nudge3), size = .6, alpha = .6) +
  geom_half_violin(data = con_pp %>% 
                     filter(instruction == 'detect' & phs_01 == 1),
                   aes(phs_01, contrast),
                   position = position_nudge(nudge3), size = .6, alpha = .6) +
  geom_point(data = con_sum %>% filter(instruction == 'detect'),
             aes(phs_01, contrast), size = 4,
             position = position_nudge(nudge3)) +
  geom_errorbar(data = con_sum %>% filter(instruction == 'detect'),
                aes(x=phs_01, ymin=contrast-ci, ymax=contrast+ci),
                position = position_nudge(nudge3),
                width = .03, size = .7) +
  geom_line(data = con_sum %>% filter(instruction == 'detect'),
            aes(phs_01, contrast), size = .75,
            position = position_nudge(nudge3)) +
  ylim(0.05,.22) +
  scale_x_continuous(breaks=c(-0.1,0.9), labels = c('No PHS', 'PHS')) +
  labs(title = 'Innocuous TSL', x = NULL, y = 'Thermal contrast') +
  theme_classic()
# noxious
ggplot() +
  geom_point(data = con_pp %>% filter(instruction == 'pain'), 
       aes(xj, contrast), alpha = .5, size = 2) +
  geom_half_violin(data = con_pp %>% 
                     filter(instruction == 'pain' & phs_01 == 0),
                   aes(phs_01, contrast),
                   position = position_nudge(nudge3), size = .6, alpha = .6) +
  geom_half_violin(data = con_pp %>% 
                     filter(instruction == 'pain' & phs_01 == 1),
                   aes(phs_01, contrast),
                   position = position_nudge(nudge3), size = .6, alpha = .6) +
  geom_point(data = con_sum %>% filter(instruction == 'pain'),
             aes(phs_01, contrast), size = 4,
             position = position_nudge(nudge3)) +
  geom_errorbar(data = con_sum %>% filter(instruction == 'pain'),
                aes(x=phs_01, ymin=contrast-ci, ymax=contrast+ci),
                position = position_nudge(nudge3),
                width = .03, size = .7) +
  geom_line(data = con_sum %>% filter(instruction == 'pain'),
            aes(phs_01, contrast), size = .75,
            position = position_nudge(nudge3)) +
  ylim(0.10,1) +
  scale_x_continuous(breaks=c(-0.1,0.9), labels = c('No PHS', 'PHS')) +
  labs(title = 'Noxious TSL', x = NULL, y = 'Thermal contrast') +
  theme_classic()
```

