---
title: "TSL2 Hypothesis 1"
author: "Alexandra G. Mitchell and Francesca Fardo"
date: "28/02/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load required packages
if (!require("pacman")) install.packages("pacman")
  pacman::p_load(DHARMa, broom.mixed, car, lme4, lmerTest, tidyverse, rcompanion, Rmisc, emmeans)
```

# Read data from cortex and select data for analysis
```{r}
# read in data
path <- getwd() #make sure set to current folder
tsl_file <- file.path("data", "anonymised_tsl2_208.csv")

# load all data file
if (file.exists(tsl_file)){
  tsl2 <- read.csv(tsl_file)
  } else {
    print('Data file does not exist, check data is in current directory.')
  }
```

# Zscore function
```{r}
calc_z <- function (value) {
  z_score <- (value - mean(value))/sd(value)
}
```

# Calculate average tsl2 thresholds, number of PHS per participant and define PHS responders
The experimental design includes two experimental factors:
(1) Instruction (2 levels: detect or pain). Participants were instructed to press a button as soon as they perceived a new thermal sensation or a new painful sensation following an auditory cue.
(2) Baseline (3 levels: 32, 38 and 44). Cooling of the skin started at 32, 38 or 44 Celsius degrees. The auditory cue was always presented at 32 Celsius degrees. 

The order of the conditions was fixed across participant and was the following
(1) task = tsl2, instruction = detect, baseline = 32 
(2) task = tsl2, instruction = detect, baseline = 38 
(3) task = tsl2, instruction = detect, baseline = 44 
(4) task = tsl2, instruction = pain, baseline = 32 
(5) task = tsl2, instruction = pain, baseline = 38 
(6) task = tsl2, instruction = pain, baseline = 44 

For each condition, participants completed 6 trials. The first 3 trials were meant to demonstrate the task. The analyses only include the last 3 trials of each condition.

```{r}
# Select trials for analysis
df1 <- tsl2 %>% 
  filter(task == "tsl2" & trials_to_keep == 1) %>% 
  dplyr::rename(subject = exp_id)

# Average TSL2 threshold (i.e. average temperature for the detection of a new thermal change or a painful sensation)
tsl2_threshold <- aggregate(df1$threshold, 
                            by = list(subject = df1$subject, 
                                      baseline = df1$baseline, 
                                      instruction = df1$instruction), 
                            FUN = mean)

# Count of paradoxical heat sensations per participants (i.e., number of cooling trials perceived as warm, hot or painfully hot)
phs_count <- aggregate(df1$phs, by = list(subject = df1$subject, 
                                         baseline = df1$baseline, 
                                         instruction = df1$instruction), 
                       FUN = sum)

# Define PHS responders (i.e., 0 = all cooling trials perceived as cold, 1 = at least one cooling trial misperceived as warm, hot or painfully hot) for each condition
phs_01 <- as.numeric(phs_count$x != 0)

# Create new dataframe for further analyses
df <- cbind(tsl2_threshold, phs = phs_count$x, phs_01)
names(df)[names(df) == 'x'] <- 'threshold' # rename column from x to avg_threshold
df$baseline <- as.numeric(df$baseline) # define baseline as double
write.csv(df, file.path("data", "summary_tsl2_208.csv"), row.names = FALSE)
df
```

# Define the total number of PHS responders and trials

```{r}
# Calculate the total number of PHS responders per condition
df_phs <- aggregate(df$phs_01, 
                    by = list(baseline = df$baseline, 
                              instruction = df$instruction), 
                    FUN = sum)
names(df_phs)[names(df_phs) == 'x'] <- 'N_phs_responders' # rename column

# Calculate the total number of PHS trials per condition
N_phs_trials <- aggregate(df$phs, 
                          by = list(baseline = df$baseline, 
                                    instruction = df$instruction), 
                          FUN = sum)

# Create new dataframe
df_phs <- cbind(df_phs, N_phs_trials = N_phs_trials$x)
df_phs
```

# Define function to perform McNemar's test and calculate effect size (Cohen's g))
Results can be double checked with an online calculator (e.g., https://www.graphpad.com/quickcalcs/McNemar1.cfm)
```{r}
phs_mcnemar_stats <- function(temp1, temp2, instr1, instr2)
{ #Subset data for the two conditions
  df1 <- subset(df, baseline == temp1 & instruction == instr1)
  df2 <- subset(df, baseline == temp2 & instruction == instr2)
  
  # Extract PHS response vector
  phs_df1 <- df1$phs_01
  phs_df2 <- df2$phs_01
  
  # Compute contingency table counts
  yy <- sum(phs_df1 == 1 & phs_df2 == 1)
  yn <- sum(phs_df1 == 1 & phs_df2 == 0)
  ny <- sum(phs_df1 == 0 & phs_df2 == 1)
  nn <- sum(phs_df1 == 0 & phs_df2 == 0)
  
  # Create a contingency table
  phs_comparison <- matrix(c(yy, yn, ny, nn),
                           nrow = 2,
                           dimnames = list("temp1" = c("phs_yes", "phs_no"),
                                           "temp2" = c("phs_yes", "phs_no")))
                      
  # Perform McNemar's test and compute effect size (Cohen's g)
  phs_stats <- mcnemar.test(phs_comparison)
  phs_cohenG <- cohenG(phs_comparison) # OR = odds ratio,  g = Cohen’s g (effect size estimates)
  
  # Return results
  results=list(phs_stats,phs_cohenG)
  return(results)
}
```

# PHS prevalence with starting temperature
Compare the number of PHS responders between conditions
```{r, warning = FALSE}
# Innocuous PHS comparisons
res1 <- phs_mcnemar_stats(38,32,"detect","detect")
res2 <- phs_mcnemar_stats(44,32,"detect","detect")
res3 <- phs_mcnemar_stats(44,38,"detect","detect")

# Noxious PHS comparisons
res4 <- phs_mcnemar_stats(38,32,"pain", "pain")
res5 <- phs_mcnemar_stats(44,32,"pain", "pain")
res6 <- phs_mcnemar_stats(44,38,"pain", "pain")

# Innocuous vs. noxious PHS comparisons
res7 <- phs_mcnemar_stats(32,32,"detect","pain")
res8 <- phs_mcnemar_stats(38,38,"detect","pain")
res9 <- phs_mcnemar_stats(44,44,"detect","pain")

# Combine the results into a data frame
results <- data.frame(
  Comparison = c("Innoc 32 vs 38", "Innoc 32 vs 44", "Innoc 38 vs 44",
                 "Nox 32 vs 38", "Nox 32 vs 44", "Nox 38 vs 44",
                 "Innoc vs Nox (32)", "Innoc vs Nox (38)", "Innoc vs Nox (44)"),
  
  McNemar_chi_squared = round(c(res1[[1]]$statistic, res2[[1]]$statistic, res3[[1]]$statistic,
                                res4[[1]]$statistic, res5[[1]]$statistic, res6[[1]]$statistic,
                                res7[[1]]$statistic, res8[[1]]$statistic, res9[[1]]$statistic), 3),
  
  McNemar_p_value = round(c(res1[[1]]$p.value, res2[[1]]$p.value, res3[[1]]$p.value,
                            res4[[1]]$p.value, res5[[1]]$p.value, res6[[1]]$p.value,
                            res7[[1]]$p.value, res8[[1]]$p.value, res9[[1]]$p.value), 3),
  
  CohensG = round(c(res1[[2]]$Global.statistics$g, res2[[2]]$Global.statistics$g, 
                   res3[[2]]$Global.statistics$g,res4[[2]]$Global.statistics$g, 
                   res5[[2]]$Global.statistics$g, res6[[2]]$Global.statistics$g,
                   res7[[2]]$Global.statistics$g, res8[[2]]$Global.statistics$g, 
                   res9[[2]]$Global.statistics$g), 3)
)

# Print the table of results
print(results)

```

# Normality assumptions of data - for modelling

# PHS rate with starting temperature
Poisson models to assess the relationship between PHS number (rate) and starting temp
```{r}
#df$phs <- as.numeric(as.character(df$phs)) #double check it is numeric
df1$baseline <- as.factor(df1$baseline)
df1$trial_z <- calc_z(df1$trial) #trial zscore
df1$subject <- as.factor(df1$subject)
df1$instruction <- as.factor(df1$instruction)


# phs by contrast and instruction
m.phs1 <- lme4::glmer(as.factor(phs) ~ baseline * instruction + (1|subject),
                     data = df1,
                     family = 'binomial')
                     #family = poisson(link = "log"))
summary(m.phs1)

or.1 <- tidy(m.phs1, conf.int=TRUE, 
               exponentiate=TRUE, effects="fixed")
or.1

# phs by contrast and instruction including trial
# trial makes very little difference here
m.phs2 <- lme4::glmer(as.factor(phs) ~ baseline * instruction + trial_z +
                       (1|subject),
                     data = df1,
                     family = 'binomial')
                     #family = poisson(link = "log"))
summary(m.phs2)

or.2 <- tidy(m.phs2, conf.int=TRUE, 
               exponentiate=TRUE, effects="fixed")
or.2

# model comparison
anova(m.phs1, m.phs2)
```
 
# Omnibus test of mixed logistic regression model
```{r}
omtest <- Anova(m.phs1, type = 2)
omtest
```


# Post hoc analysis of regression model
```{r}
phoc <- emmeans(m.phs1, ~ baseline * instruction)
phoc

pw <- pairs(phoc, simple = 'baseline')
pw

```

